---
title: "Non Linear Models"
output: html_notebook
---

#Polynomial Regression
Use poly() command within the lm function.

poly(age,4) returns a matrix whose columns are a basis of orthogonal polynomials (each column is a linear combination of the variables age, age^2, age^3 and age^4).

We can use poly() to obtain age, age^2, age^3 and age^4 directly by using the raw=TRUE argument. 

Even though it affects the coefficient estimates, it does not affect the fitted values. 

```{r}
library(ISLR)

fit=lm(wage???poly(age,4) , data=Wage)
coef(summary(fit))

fit2=lm(wage???poly(age ,4,raw=T),data=Wage)
coef(summary(fit2))

agelims=range(age) #a vector of two elements (min and max)
age.grid=seq(from=agelims[1], to=agelims[2]) #by=1 by default
preds=predict(fit, newdata=list(age=age.grid), se=TRUE)
preds2=predict(fit2, newdata=list(age=age.grid), se=TRUE)
max(abs(preds$fit-preds2$fit)) #the maximum difference between the fitted values of fit and fit2
```

There are several other equivalent ways of fitting polynomial regression model.

We can use the wrapper function I() to protect terms since the ^ symbol has a special meaning in formulas.

cbind() function does the same more compactly by building a matrix from a collection of vectors.

```{r}
fit2a=lm(wage???age+I(age^2)+I(age^3)+I(age^4),data=Wage)
coef(summary(fit2a))

fit2b=lm(wage???cbind(age, age^2, age^3, age ^4),data=Wage)
coef(summary(fit2b))
```

We use our model to predict wages using age and plot the data (dark grey) together with the fit (blue line with confidence interval). 

```{r}
preds=predict(fit, newdata=list(age=age.grid), se=TRUE)
se.bands=cbind(preds$fit+2*preds$se.fit, preds$fit-2*preds$se.fit) #create the confidence bands

par(mfrow=c(1,2), mar=c(4.5,4.5,1,1), oma=c(0,0,4,0)) #mar and oma to control the margins of the plot
plot(age, wage, xlim=agelims, cex =.5, col="darkgrey") #scatterplot at the background
title("Degree-4 Polynomial",outer=T)
lines(age.grid, preds$fit, lwd=2, col="blue")
matlines(age.grid, se.bands, lwd=1, col="blue", lty=3)
```

We need to decide on the degree of the polynomial to use. One way to do this is by using hypothesis tests.

We fit models ranging from linear to a degree-5 polynomial and seek to determine the simplest model which is sufficient to explain the relationship.

We use the anova() function, which performs a F-test in order to test the null hypothesis that a model M1 is sufficient to explain the data against the alternative hypothesis that a more complex model M2 is required.

In order to use the anova() function, M1 and M2 must be nested models: the predictors in M1 must be a subset of the predictors in M2.

The p-value comparing the linear model M1 to the quadratic model M2 is essentially zero (2.2e-16), indicating that a linear fit is not sufficient.

The p-value comparing the cubic and degree-4 polynomials is approximately 5% while the degree-5 polynomial seems unnecessary
because its p-value is 0.37.

Hence, either a cubic or a degree-4 polynomial appear to provide a reasonable fit to the data.

```{r}
fit.1=lm(wage???age, data=Wage)
fit.2=lm(wage???poly(age,2), data=Wage)
fit.3=lm(wage???poly(age,3), data=Wage)
fit.4=lm(wage???poly(age,4), data=Wage)
fit.5=lm(wage???poly(age,5), data=Wage)
anova(fit.1, fit.2, fit.3, fit.4, fit.5)
```

We can also use anova() when we have other terms in the model.

```{r}
fit.1=lm(wage???education+age, data=Wage)
fit.2=lm(wage???education+poly(age,2), data=Wage)
fit.3=lm(wage???education+poly(age,3), data=Wage)
anova(fit.1, fit.2, fit.3)
```

##Logistic Regression
We want to predict whether an individual earns more than $250,000 per year.

First we create the appropriate response vector. We use the wrapper I(wage>250) to create a binary response variable. The expression "wage>250" evaluates to a logical variable containing TRUEs and FALSEs, which glm() coerces to binary by setting the TRUEs to 1 and the FALSEs to 0.

Then we apply the glm() function using family="binomial" in order to fit a polynomial logistic regression model.

We use our model to predict probability using age. The default prediction type for a glm() model is type="link". This means we get predictions and standard erros for the logit(log odds): log(p/1-p)=X??

In order to obtain confidence intervals for probability, we use the transformation p=exp(X??)/(1+exp(X??)).

Note that we could have directly computed the probabilities by selecting type="response" option in the predict() function. However, the corresponding confidence intervals would not have been sensible because we would end up with negative probabilities!

```{r}
fit=glm(I(wage>250)???poly(age,4), data=Wage, family=binomial)

preds=predict(fit, newdata=list(age=age.grid), se=T)
pfit=exp(preds$fit)/(1+exp(preds$fit)) #transformation
se.bands.logit=cbind(preds$fit+2*preds$se.fit, preds$fit-2*preds$se.fit)
se.bands=exp(se.bands.logit)/(1+exp(se.bands.logit)) #transformation
```

We plot the data together with the fit.

We have drawn the age values corresponding to the observations with wage values above 250 as gray marks on the top of the plot, and those with wage values below 250 are shown as gray marks on the bottom of the plot. 

We used the jitter() function to jitter the age values a bit so that observations with the same age value do not cover each other up.

```{r}
plot(age, I(wage>250), xlim=agelims, type="n", ylim=c(0,.2))
points(jitter(age), I((wage>250)/5), cex=.5, pch ="|", col="darkgrey") #(wage>250)/5 to convert 1 to 0.20 so it would be shown at the top of the plot
lines(age.grid, pfit,lwd=2, col="blue")
matlines(age.grid, se.bands, lwd=1, col="blue", lty=3)
```

#Step function
Use the cut() function to fit a step function

cut() automatically picked the cutpoints at 33.5, 49, and 64.5 years of age. We could also have specified our own cutpoints directly using the breaks option.

The function cut() assigns each observation to one of the 4 intervals, the lm() function then creates a set of dummy variables for use in the regression.

The age<33.5 category is left out, so the intercept coefficient of $94,160 can be interpreted as the average salary for those under 33.5 years of age. 

```{r}
table(cut(age,4)) #4 intervals so 3 cutpoints 
fit=lm(wage???cut(age,4), data=Wage)
coef(summary(fit))
```

#Splines
Regression splines can be fit by constructing an appropriate matrix of basis functions.

The bs() function in the the splines library generates the entire matrix of basis functions for splines with specified set of knots. 

We prespecify the knots at ages 25, 40, and 60. This produces a spline with six basis functions. Also, recall that a cubic spline with three knots has seven degrees of freedom.

We could also use the df option to produce a spline with knots at uniform quantiles of the data. 

By default, cubic splines are produced. However, the function bs() also has a degree argument, so we can fit splines of any degree, rather than the default degree of 3.

```{r}
library(splines)
dim(bs(age, knots=c(25,40,60))) #prespecified knots 
dim(bs(age, df=6)) #prespecified degree of freedom produces knots at uniform quantiles
attr(bs(age, df=6), "knots")

fit=lm(wage???bs(age, knots=c(25,40,60)), data=Wage) 

pred=predict(fit, newdata=list(age=age.grid), se=T)
plot(age, wage, col="gray")
lines(age.grid, pred$fit ,lwd=2)
lines(age.grid, pred$fit+2*pred$se, lty="dashed")
lines(age.grid, pred$fit-2*pred$se, lty="dashed")
```

#Natural Spline
We can use the ns() function to fit a natural spline with four degrees of freedom.

As with the bs() function, we could instead specify the knots directly using the knots option.

```{r}
fit2=lm(wage???ns(age, df=4), data=Wage)
pred2=predict(fit2, newdata=list(age=age.grid), se=T)

plot(age, wage, col="gray")
lines(age.grid, pred2$fit, col="red", lwd=2)
```

#Smoothing Spline
We use the smooth.spline() function to fit a smoothing spline.

In the first call to smooth.spline(), we specified df=16. The function then determines which value of ?? leads to 16 degrees of freedom. 

In the second call to smooth.spline(), we select the smoothness level by cross-validation. This results in a value of ?? that yields 6.8 degrees of freedom.

```{r}
fit=smooth.spline(age, wage, df=16)
fit2=smooth.spline(age, wage, cv=TRUE)
fit2$df

plot(age, wage, xlim=agelims, cex =.5, col="darkgrey")
title("Smoothing Spline")
lines(fit, col="red", lwd =2)
lines(fit2, col="blue", lwd=2)
legend("topright", legend=c("16 DF" ,"6.8 DF"), col=c("red","blue"), lty=1, lwd=2, cex =.8)
```

#Local Regression
We use the loess() function to perform local regression.

We perform local linear regression using spans of 0.2 and 0.5: that is, each neighborhood consists of 20 % or 50 % of the observations. The larger the span, the smoother the fit. 

The locfit library can also be used for fitting local regression models.

```{r}
fit=loess(wage???age, span=.2, data=Wage)
fit2=loess(wage???age, span=.5, data=Wage)

plot(age, wage, xlim=agelims, cex =.5, col="darkgrey")
title("Local Regression")
lines(age.grid, predict(fit,data.frame(age=age.grid)), col="red", lwd=2)
lines(age.grid, predict(fit2,data.frame(age=age.grid)), col="blue", lwd=2)
legend("topright", legend=c("Span=0.2","Span=0.5"), col=c("red","blue"), lty=1, lwd=2, cex=.8)
```

#GAMs
We first fit a GAM to predict wage using natural spline functions of year and age, treating education as a qualitative predictor. We then fit the model using smoothing splines rather than natural splines.

We use the gam() function in gam library.The s() function, which is part of the gam library, is used to indicate that we would like to use a smoothing spline. 

The generic plot() function recognizes that gam.m3 is an object of class gam, and invokes the appropriate plot.gam() method. 

```{r}
library(gam)
gam1=gam(wage???ns(year,4)+ns(age,5)+education, data=Wage) #4 df for year and 5 df for age
gam2=gam(wage???s(year,4)+s(age,5)+education, data=Wage)

par(mfrow=c(1,3))
plot(gam2, se=TRUE, col="blue") #setting se=TRUE will produce the confidence bands on the graph
plot(gam1, se=TRUE, col="red") 
```

In these plots, the function of year looks rather linear. We can perform a series of ANOVA tests in order to determine which of these three models is best: a GAM that excludes year (M1), a GAM that uses a linear function of year (M2), or a GAM that uses a smoothing spline function of year (M3).

We find that there is compelling evidence that a GAM with a linear function of year is better than a GAM that does not include year at all (p-value = 0.00014). 

However, there is no evidence that a non-linear function of year is needed (p-value = 0.349). Therefore, based on the results of this ANOVA, M2 is preferred. 

```{r}
gam.m1=gam(wage???s(age,5)+education, data=Wage)
gam.m2=gam(wage???year+s(age,5)+education, data=Wage)
gam.m3=gam(wage???s(year,4)+s(age,5)+education, data=Wage)
anova(gam.m1, gam.m2, gam.m3, test="F")
```

The summary() function produces a summary of the gam fit.

For Anova for Nonparametric Effects, the p-values for year and age correspond to a null hypothesis of a linear relationship versus the alternative of a non-linear relationship. 

The large p-value for year reinforces our conclusion from the ANOVA test that a linear function is adequate for this term. 

However, there is very clear evidence that a non-linear term is required for age. 

```{r}
summary(gam.m3)
```

We can make predictions from gam objects using the predict() method (just like for lm objects). 

```{r}
preds=predict(gam.m2, newdata=Wage)
```

We can use local regression fits as building blocks in a GAM, using the lo() function. 

We can also use the lo() function to create interactions before calling the gam() function.

We can plot the resulting two-dimensional surface by installing the akima package. 

```{r}
gam.lo=gam(wage???s(year,df=4)+lo(age,span =0.7)+education, data=Wage) #local regression for the age term, with a span of 0.7
plot(gam.lo, se=TRUE, col="green")

gam.lo.i=gam(wage???lo(year,age,span=0.5)+education, data=Wage) #an interaction term between year and age, fit by a local regression surface
library(akima)
plot(gam.lo.i)
```

##Logsitic Regression
We use the I() function in constructing the binary response variable, and set family=binomial.

```{r}
gam.lr=gam(I(wage>250)???year+s(age,df=5)+education, family=binomial, data=Wage)
par(mfrow=c(1,3))
plot(gam.lr, se=T, col="green") #note the wide confidence interval for the first category of education
```

Since there are no high earners in the <HS category, we fit a logistic regression GAM using all but this category.

```{r}
table(education, I(wage>250)) #education and I(wage>250) are objects which can be interpreted as factors so R build a table of the counts at each combination of factor levels

gam.lr.s=gam(I(wage>250)???year+s(age,df=5)+education, family=binomial, data=Wage, subset=(education != "1. < HS Grad"))
par(mfrow=c(1,3))
plot(gam.lr.s, se=T, col="green")
```

